Superphy: Real-time prediction of \textit{E. coli} phenotype using graph-based genome sequence storage and retrieval with the \textit{spfy} module

### 2017-11-27 paper-webserver.tex
EDIT: some restructuring of the comparison section

% para
% focus on application
The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns {ronholm2016navigating}, or Polymerase Chain Reaction (PCR)-based amplication to identify known VFs or AMR genes, along with other tests, to characterize bacterial isolates in outbreak and surveillance settings.

@@

\subsection{Comparison with other bioinformatic pipeline technologies}

% namely galaxy
Other scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP are used to run analysis modules on WGS data.
Galaxy aims to provide a reproducible compuation-based research environment which is accessible to individuals without programming knowledge.
TODO: BAP and IRIDA, since they were the examples from above
IRIDA...
BAP...

The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, is the closest analogue to Spfy, and provides an integrated analysis pipeline for bacterial WGS data as a web service.
BAP is novel in its combined approach to genome analysis as different programs, such as for VF and AMR determination, are included by default into the pipeline; Spfy provides similar functionallity with an expanded focus on integrated result storage and big-data analyses.
In place of a MySQL database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistant graph database.
This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.

>>>>

% para
% focus on application
The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns, or Polymerase Chain Reaction (PCR)-based amplication to identify known VFs or AMR genes, along with other tests \cite{ronholm2016navigating}, to characterize bacterial isolates in outbreak and surveillance settings.

@@

\subsection{Comparison with other bioinformatic pipeline technologies}
% trying to cut down on details that would be better suited for the "Implemntation" section.

% namely galaxy
Other scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP are used to run analysis modules on WGS data.
Galaxy aims to provide a reproducible compuation-based research environment which is accessible to individuals without programming knowledge. Galaxy tackles the problem of linking different analysis software together, by defining interdependencies using a custom schema. A visual workflow editor is also provided for ease of use.
IRIDA is build on top of the Galaxy framework, and adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage. IRIDA takes a project-based approach, with sequences stored per project, and results store linearly per sequence. Focus is given on refined controls for collaborating on projects, and common terms found in the GenEpiO ontology are used to describe results.

The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark provides an integrated analysis pipeline for bacterial WGS data as a web service.
BAP takes a combined approach to genome analysis including different programs, such as for VF and AMR determination, by default into the pipeline.
A record of the different files generated by an analysis are stored in a relational database, and the record is queried to determine which files are returned to the user \cite{thomsen2016bacterial}.

In Spfy, we take a more opionionated approach to result storage and data integration. The benefits of real-time analysis pipelines as found in BAP, and the use of the GenEpiO ontology as in IRIDA, are included in Spfy.
On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
In place of a relational database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistant graph database.
This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses. In all, Spfy provides similar functionallity with an expanded focus on integrated result storage and big-data analyses.
